Active Learning from Blackbox to Timed Connectors

### 2016-04-04 paper/paper.tex
.

Distributed real-time embedded systems (DRES) are reforming our lives with the
Internet of Things(IoT), wherein individual components are composed via \emph{connectors} 
to build complex systems. Such systems could be distributed logically or physically, which
makes coordination processes even more complicated. In this case, we need to specify the coordination
processes with \emph{coordination languages}, such that formal techniques can be applied to
guarantee their reliability.

\emph{Timed Reo} is a real-time extension of the coordination language \emph{Reo}, which can be used
to describe the coordination process in DRES clearly and intuitively. Different formal
semantics have been proposed to specify the behavior of timed Reo.
For example, an operational semantics based on \emph{timed constraint automata} was raised by Baier et
al.\cite{DBLP:conf/sefm/ArbabBBR04}, where a variant of LTL was also proposed to describe the
properties of timed connectors.
In \cite{DBLP:conf/tase/Meng12}, a UTP design model was proposed to specify the behavior of timed
connectors..

Formal verification and validation techniques have also been proved applicable for timed connectors, 
e.g. a comformance testing framework on timed connectors has been proposed in \cite{DBLP:conf/tase/LiCWS15}.
A SAT-based bounded model checking approach was adapted in \cite{DBLP:journals/scp/Kemper12} for
verifying connectors. Although these solutions are practical and impressive, there is one important
question faced by most of them: \emph{how to obtain the formal models?}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.4\textwidth]{./images/howto.png}
  \end{center}
  \caption{The Active Learning Framework}
  \label{fig:howto}
\end{figure}

@@

The rest of the paper is organized as follows: In Section
\ref{sec:preliminaries} we briefly illustrate some basic concepts, including the coordination
language Reo, Mealy machine, and active automata learning. Section \ref{sec:semantics} defines an
operational semantics of Reo, which is used in Section \ref{sec:activelearning} to show how to
extract Reo models from blackboxes by means of active learning. In Section
\ref{sec:experiment}, we discuss the implementation and optimization of the approach in
\texttt{Golang}. Finally, Section VI concludes the paper.

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed through its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always ready to accept data items. These items will be
    dispensed through its sink end simultaneously if possible, otherwise they will be lost.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to be accepted
    simultaneously through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty. 
  \item [-] A t-timer channel accepts any data item from its source end, and later dispense it to
    its sink end after a delay of $t$ time units.
    \footnote{
      Here the interpretation of t-Timer is a bit different as in other works like
      \cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/tase/Meng12}, but we can easily transform between the
      two kinds of behavior by composing with some other basic channels.
      }
\end{itemize}

@@

Components can be linked to source nodes or sink nodes. A component can write data items to its
corresponding source node only if the data item can be dispensed simultaneously to all source ends
on this node. Similarly, a component can read a data item only if there is at least one readable
sink end on its corresponding node. A mixed node non-determinstically selects and takes a data item
from one of its coincident sink ends and replicates it into all of its coincident source ends.

@@

Active learning \cite{settles2010active} is a special case of semi-supervised machine learning where
a learning algorithm is able to interactively query the target systems to obtain the desired outputs
under certain inputs. By well-designed query strategy, active learning is able to obtain more
accurate models with smaller dataset.

@@

\subsection{Time Domain}
Time has been investigated in several extensions of Reo. For example, timed
Reo\cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/fmoods/MengA07}, hybrid Reo\cite{DBLP:conf/icfem/ChenSS14}, etc.
Generally, these models are designed to handle real-time behavior where time is defined on the
real number field $\mathbb{R}$. However, there are also some works like
\cite{DBLP:journals/fmsd/PrabhakarDM015} where
rational time indeed  makes things easier. In this paper, we choose the rational number field
$\mathbb{Q}$ as our time domain, which simplifies discretization of timed behavior greatly.

@@

Now we use \emph{access sequence} to describe the states. Given an input sequence $s$, the access
sequence $acc(s,D)$ of $s$ is defined as the shortest sequence in $\{s'\in\mathcal{I}^*|s'\sim_D
s\}$\footnote{When there are multiple satisfying sequences with the same length, we use a certain tactic
to pick up one from them determinstically.}.

@@

Now we consider the reason that leads to existance of counter examples. Since
the suffix set $D$ is used to distinguish states, the existance of counter-examples
shows that the current suffix set $D$ is not powerful enough to recognize the uncovered states in
$next$ (see Algorithm \ref{alg:buildtable}).

@@

Both \emph{Reo Coordination Models} and \emph{Adapted L* Algorithm} are implemented in
Golang\cite{golang}, which is widely known for its elegant design and impressive efficiency.
Moreover, with a CSP-style\cite{DBLP:books/ph/Hoare85} cocurrency model, \texttt{Golang} shares many
similar ideas with Reo, and in turn makes our implementation much more natural.

@@

If previous simulation results are stored in a well-maintained cache, the time-cost in
simulation process could be reduced signficiantly. In this work, we use a multiway tree to buffer
these results.

@@

\section{Conclusion and Future Work}
In this paper, we come up with an approach to extract timed connectors from blackbox
implementations. We propose \emph{parameterized Mealy machine} to describe the behavior of Reo
connectors. Then we define the product operator and link operator, which can be used to
construct complex connectors from simpler ones.
Besides, we show how PMMs behave as the bridge between Reo connectors and concrete Mealy
machines with timed action T provided.

We also adapt the well-known active learning algorithm L* to deal with time domain.
When time is taken into consideration, the original L* algorithm run into a performance bottleneck.
By a tree-style cache, we make significant reduction on membership queries and, in turn,
improve the performance of learning algorithm. As a by-product, we also encapsulate the Reo
connector as a distributable package which could contribute to concurrent programming.

>>>>

Distributed real-time embedded systems (DRES) are reforming our lives with the
Internet of Things, wherein individual components are composed via \emph{connectors}
to build complex systems. Such systems could be distributed logically or physically, which
makes coordination processes even more complicated. In this case, we need to specify the coordination
processes with \emph{coordination languages}, such that formal techniques can be applied to
guarantee their reliability.

\emph{Timed Reo} \cite{DBLP:conf/sefm/ArbabBBR04} is a real-time extension of the coordination language \emph{Reo}
\cite{DBLP:journals/mscs/Arbab04}, which can be used
to describe the coordination process in DRES clearly and intuitively. Different formal
semantics have been proposed to specify the behavior of timed Reo.
For example, an operational semantics based on \emph{timed constraint automata} was proposed in 
\cite{DBLP:conf/sefm/ArbabBBR04}, where a variant of LTL was also proposed to describe the
properties of timed connectors.
In \cite{DBLP:conf/tase/Meng12}, a UTP design model was proposed to specify the behavior of timed
connectors.

Formal verification and validation techniques have also been proved applicable for timed connectors,
e.g. a conformance testing framework on timed connectors has been proposed in \cite{DBLP:conf/tase/LiCWS15}.
A SAT-based bounded approach was adapted in \cite{DBLP:journals/scp/Kemper12} for model checking
connectors. Although these solutions are practical and impressive, there is one important
question faced by most of them: \emph{how to obtain the formal models?}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.45\textwidth]{./images/howto.png}
  \end{center}
  \caption{The Active Learning Framework}
  \label{fig:howto}
\end{figure}

@@

The rest of the paper is organized as follows: In Section
\ref{sec:preliminaries} we illustrate some basic concepts, including the coordination
language Reo, Mealy machine, and active automata learning. Section \ref{sec:semantics} briefly presents an
operational semantics of Reo based on PMM, which is used in Section \ref{sec:activelearning} to show how to
extract Reo models from blackboxes by means of active learning. In Section
\ref{sec:experiment}, we discuss the implementation and optimization of the approach in
\texttt{Golang}. Finally, Section VI concludes the paper.

@@

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed through its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always ready to accept data items. These items will be
    dispensed through its sink end simultaneously if possible, otherwise they will be lost.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to be accepted
    simultaneously through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asynchronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty.
  \item [-] A t-timer channel accepts any data item from its source end, and later dispense it to
    its sink end after a delay of $t$ time units.
    \footnote{
      Here the interpretation of t-Timer is a bit different as in other works like
      \cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/tase/Meng12}, but we can easily transform between the
      two kinds of behavior by composing with some other basic channels.
      }
\end{itemize}

@@

Components can be linked to source nodes or sink nodes. A component can write data items to its
corresponding source node only if the data item can be dispensed simultaneously to all source ends
on this node. Similarly, a component can read a data item only if there is at least one readable
sink end on its corresponding node. A mixed node non-deterministically selects and takes a data item
from one of its coincident sink ends and replicates it into all of its coincident source ends.

@@

Active learning \cite{settles2010active} is a special case of semi-supervised machine learning where
a learning algorithm is able to interactively query the target systems to obtain the desired outputs
under certain inputs. By well-designed query strategy, active learning is able to obtain more
accurate models with smaller data-set.

@@

\subsection{Time Domain}
Time has been investigated in several extensions of Reo. For example, timed
Reo\cite{DBLP:conf/sefm/ArbabBBR04}, resource-sensitive timed Reo\cite{DBLP:conf/fmoods/MengA07}, hybrid Reo\cite{DBLP:conf/icfem/ChenSS14}, etc.
Generally, these models are designed to handle real-time behavior where time is defined on the
real number field $\mathbb{R}$. However, there are also some works like
\cite{DBLP:journals/fmsd/PrabhakarDM015} where
rational time indeed  makes things easier. In this paper, we choose the rational number field
$\mathbb{Q}$ as our time domain, which simplifies discretization of timed behavior greatly.

@@

Now we use \emph{access sequence} to describe the states. Given an input sequence $s$, the access
sequence $acc(s,D)$ of $s$ is defined as the shortest sequence in $\{s'\in\mathcal{I}^*|s'\sim_D
s\}$\footnote{When there are multiple satisfying sequences with the same length, we use a certain tactic
to pick up one from them deterministically.}.

@@

Now we consider the reason that leads to existence of counter examples. Since
the suffix set $D$ is used to distinguish states, the existence of counter-examples
shows that the current suffix set $D$ is not powerful enough to recognize the uncovered states in
$next$ (see Algorithm \ref{alg:buildtable}).

@@

Both \emph{Reo Coordination Models} and \emph{Adapted L* Algorithm} are implemented in
Golang\cite{golang}, which is widely known for its elegant design and impressive efficiency.
Moreover, with a CSP-style\cite{DBLP:books/ph/Hoare85} concurrency model, \texttt{Golang} shares many
similar ideas with Reo, and in turn makes our implementation much more natural.

@@

If previous simulation results are stored in a well-maintained cache, the time-cost in
simulation process could be reduced significantly. In this work, we use a multiway tree to buffer
these results.

@@

\section{Conclusion and Future Work}
In this paper, we come up with an approach to extract timed connectors from blackbox
implementations. We propose \emph{parameterized Mealy machine} to describe the behavior of Reo
connectors. Then we define the product operator and link operator, which can be used to
construct complex connectors from simpler ones.
Besides, we show how PMMs behave as the bridge between Reo connectors and concrete Mealy
machines with timed action $T$ provided.

We also adapt the well-known active learning algorithm $L*$ to deal with time domain.
When time is taken into consideration, the original $L*$ algorithm run into a performance bottleneck.
By a tree-style cache, we make significant reduction on membership queries and, in turn,
improve the performance of learning algorithm. As a by-product, we also encapsulate the Reo
connector as a distributable package which could contribute to concurrent programming.

### 2016-04-04 shortpaper/paper.tex
.

\begin{document}
\maketitle 
\begin{abstract}
  Coordination models and languages play a key role in formally specifying the communication and
  interaction among different components in large-scale distributed and concurrent systems. In this
  paper, we use active learning to extract timed connector models from black-box
  system implementation. 
  We first introduce parameterized Mealy machine(PMM) as an operational semantic
  model for channel-based coordination language Reo. PMM serves as a bridge
  between Reo connectors and Mealy machines. With the product operator, complex connectors can be
  constructed by joining basic channels and the PMMs of connectors can be transformed into Mealy
  machines. Moreover, we adapt L*, a well-known learning algorithm, to timed connectors (in the form
  of Mealy machines). The new algorithm has shown its efficiency in multiple case studies. 
\end{abstract}

@@

Distributed real-time embedded systems (DRES) are reforming our lives with the
Internet of Things(IoT), wherein individual components are composed via \emph{connectors} 
to build complex systems. Such systems could be distributed logically or physically, which
makes coordination processes even more complicated. In this case, we need to specify the coordination
processes with \emph{coordination languages}, such that formal techniques can be applied to
guarantee their reliability.

\emph{Timed Reo} is a real-time extension of the coordination language \emph{Reo}, which can be used
to describe the coordination process in DRES clearly and intuitively. 
\todo{describe semantics in a more short way}
%Different formal semantics have been proposed to specify the behavior of timed Reo.
%For example, an operational semantics based on \emph{timed constraint automata} was raised by Baier et
%al.\cite{DBLP:conf/sefm/ArbabBBR04}, where a variant of LTL was also proposed to describe the
%properties of timed connectors.
%In \cite{DBLP:conf/tase/Meng12}, a UTP design model was proposed to specify the behavior of timed
%connectors..

Formal verification and validation techniques have also been proved applicable for timed connectors, 
e.g. a comformance testing framework on timed connectors has been proposed in \cite{DBLP:conf/tase/LiCWS15}.
A SAT-based bounded model checking approach was adapted in \cite{DBLP:journals/scp/Kemper12} for
verifying connectors. Although these solutions are practical and impressive, there is one important
question faced by most of them: \emph{how to obtain the formal models?}

\emph{Correctness} of connectors is closely related to some low-level implementation details.
For example, well-written code may behave dramatically weird with an improper set of concurrency
primitives. Such scenarios happen frequently in embedded systems with different hardware platforms or
operating systems. Consequently, manually extracting a proper model from an existing connector
implementation seems rather unreliable, even with reference to its source code.

As a branch of \emph{machine learning} technique, active learning offers a means to obtain
high-level models from low-level models. Works in \cite{de2010grammatical,
DBLP:journals/iandc/Angluin87, DBLP:conf/fase/RaffeltS06} show interesting examples where active
learning is used to extract \emph{Mealy machines} or \emph{regular languages} without time domain.

%\begin{figure}[ht]
  %\begin{center}
    %\includegraphics[width=.4\textwidth]{./images/howto.png}
  %\end{center}
  %\caption{The Active Learning Framework}
  %\label{fig:howto}
%\end{figure}

In this paper we address this question by proposing an active learning framework as shown in Fig.
\ref{fig:howto} to automatically extract timed connector models from blackbox implementations. 
To achieve this goal, we first introduce \emph{parameterized Mealy machine} as a parameterized
semantics for timed Reo connectors, which can be transformed to concrete Mealy machines with a
given alphabet. Then the $L*$ algorithm \cite{DBLP:journals/iandc/Angluin87} is adapted and
optimized to extract Mealy machines with timed action from the connector implementations as
blackboxes.

\todo{use a short paragraph to describe the preliminaries}

The rest of the paper is organized as follows: In Section
\ref{sec:preliminaries} we briefly illustrate some basic concepts, including the coordination
language Reo, Mealy machine, and active automata learning. Section \ref{sec:semantics} defines an
operational semantics of Reo, which is used in Section \ref{sec:activelearning} to show how to
extract Reo models from blackboxes by means of active learning. In Section
\ref{sec:experiment}, we discuss the implementation and optimization of the approach in
\texttt{Golang}. Finally, Section VI concludes the paper.

@@

Mealy machines consist of \emph{states} and \emph{transitions}.
As for blackboxes, input sequence $s\in\mathcal{I}^*$ can be used to denote its state 
after executing $s$. Similarly, for all $a\in\mathcal{I}$, we use $sa$ to denote the successor
state of $s$ after executing $a$.
However, under such notations, there are \emph{infinite} number of states, among which most are
reduplicate.

>>>>

\begin{document}
\maketitle 
\begin{abstract}
  Coordination models and languages play a key role in formally specifying the communication and
  interaction among different components in large-scale concurrent systems. In this
  paper, we use active learning to extract timed connector models from black-box
  system implementation. 
  Firstly, parameterized Mealy machines(PMM) is introduced as an operational semantic
  model for channel-based coordination language Reo. With product and link operators defined, we can
  construct complex connectors by joining basic channels in form of PMM. Moreover, with a
  concretize mapping function, PMMs can be easily transformed into Mealy machines, and the latter can be
  extracted by an optimized L* algorithm.
\end{abstract}

@@

Distributed real-time embedded systems (DRES) are reforming our lives with the
Internet of Things(IoT), wherein individual components are composed via \emph{connectors} 
to build complex systems. 
%Such systems could be distributed logically or physically, which
%makes coordination processes even more complicated. 
In this case, we need to specify the coordination
processes with \emph{coordination languages}, such that formal techniques can be applied to
guarantee their reliability.

As a real-time extension of \emph{Reo}, \emph{Timed
Reo}\cite{DBLP:conf/sefm/ArbabBBR04,DBLP:conf/tase/Meng12} is used in this paper to formalize
connectors in DRES. 
Reo is a channel based exogenous coordination language proposed by F. Arbab in
\cite{DBLP:journals/mscs/Arbab04}, which is a powerful ``glue language'' in component-based
development\cite{DBLP:journals/sigsoft/Gill03}. What's more, 
Formal verification and validation techniques have also been proved applicable for timed connectors, 
e.g. a comformance testing framework on timed connectors has been proposed in \cite{DBLP:conf/tase/LiCWS15}.
A SAT-based bounded model checking approach was adapted in \cite{DBLP:journals/scp/Kemper12} for
verifying connectors.

However, most of these tools heavily rely on manually modelling.
\emph{Correctness} of connectors is closely related to some low-level implementation details.
For example, well-written code may behave dramatically weird with an improper set of concurrency
primitives. Such scenarios happen frequently in embedded systems with different hardware platforms or
operating systems. Consequently, manually extracting a proper model from an existing connector
implementation seems rather unreliable, even with reference to its source code.

In this paper we address this question by means of active learning \cite{de2010grammatical,
DBLP:journals/iandc/Angluin87, DBLP:conf/fase/RaffeltS06}.
First of all, we introduce \emph{parameterized Mealy machine} as a parameterized
semantics for timed Reo, which can be transformed to concrete Mealy machines with a
specified alphabet. Then the $L*$ algorithm \cite{DBLP:journals/iandc/Angluin87} is adapted and
optimized to extract Mealy machines with timed action from the connector implementations as
blackboxes.

The rest of the paper is organized as follows: In Section
Section \ref{sec:semantics}, we defines an
operational semantics of Reo, which is used in Section \ref{sec:activelearning} to show how to
extract Reo models from blackboxes by means of active learning. In Section
\ref{sec:experiment}, we discuss the implementation and optimization of the approach. 
Finally, Section VI concludes the paper. Due to the limit of space, many technical details are
omitted in this paper.
Please refer to \todo{github repo} for further contents.

@@

Mealy machines consist of \emph{states} and \emph{transitions}.
As for blackboxes, input sequence $s\in\mathcal{I}^*$ can be used to denote its state 
after executing $s$. Similarly, for all $a\in\mathcal{I}$, we use $sa$ to denote the successor
state of $s$ after executing $a$.
However, under such notations, there are \emph{infinite} number of states, among which most are
duplicate.
